SDSS-IV MaNGA: The Universal Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-08-19 maintex.tex
minor edit to abstract

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 220 interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxies is matched to a set of 20 control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. We describe the profiles of the $\Delta$log(sSFR) enhancement as a universal profile because the general shape of the profile is unchanged by parameters like stellar mass, mass ratio, or projected separation. The shape of the profile is a central peak in $\Delta$log(sSFR) which linearly falls to zero around 1.0 R/\reff. Parameters like the mass ratio or projected separation may scale the height of the profile's central peak, but the general shape of the profile is preserved. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate of 1$-$6\%, assuming a constant burst merger induced star formation over a 1 Gyr period. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

>>>>

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 220 interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxy is matched to a set of 20 control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. We describe the profiles of the $\Delta$log(sSFR) enhancement as a universal profile because the general shape of the profile is unchanged by parameters like stellar mass, mass ratio, or projected separation. The shape of the profile is a central peak in $\Delta$log(sSFR) which linearly falls to zero around 1.0 R/\reff. Parameters like the mass ratio or projected separation may scale the height of the profile's central peak, but the general shape of the profile is preserved. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate of 1$-$6\%, assuming a constant burst merger induced star formation over a 1 Gyr period. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.
